---
title: "Investigating TAD boundaries"
author: "Spiro Stilianoudakis"
date: "April 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(dplyr)

setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data_analysis/data")
arrowhead_data <- read.table("arrowhead_data.txt", header=T)
dim(arrowhead_data)
```

# There are 9274 TAD boundary intervals in the data

# Plotting TAD boundaries for Chromosome 22

```{r echo=FALSE}

arrowhead_data <- arrowhead_data[,1:3]
plot(arrowhead_data$Start[which(arrowhead_data$Chromosome=="chr22")],arrowhead_data$End[which(arrowhead_data$Chromosome=="chr22")],pch='.', col="blue", xlab="Start", ylab="End")
```

# Zooming in to a square window of 32000000-34000000 coordinates

```{r echo=FALSE}
plot(arrowhead_data$Start[which(arrowhead_data$Chromosome=="chr22")],arrowhead_data$End[which(arrowhead_data$Chromosome=="chr22")], col="blue", xlim=c(3.2e+07,3.4e+07),ylim=c(3.2e+07,3.4e+07), xlab="Start", ylab="End")
#adding diagonal line
abline(0,1)

arrowhead_data[which(arrowhead_data$Start>=32000000 & arrowhead_data$Chromosome=="chr22" & arrowhead_data$End<=34000000),]

# There appear to be situations where we have a TAD within a TAD
# In other words, the starting coordinate is the same for two TAD boundaries
```

# How many ways can TAD boundaries share coordinates?

## 1. They could share the same Start site

Example:

| CHR | Start    | End      |
|:---:|:--------:|:--------:|
| 22  | 32370000 | 32890000 |
| 22  | 32370000 | 32495000 |

## 2. They could share the same End site

Example:

| CHR | Start    | End      |
|:---:|:--------:|:--------:|
| 22  | 33695000 | 33985000 |
| 22  | 33900000 | 33985000 |

## 3. They could be consecutive (start site = end site)

Example:

| CHR | Start    | End      |
|:---:|:--------:|:--------:|
| 22  | 33605000 | 33695000 |
| 22  | 33695000 | 33985000 |

## 4. They could be the same start and end site (duplicate intervals)

Example:

| CHR | Start    | End      |
|:---:|:--------:|:--------:|
| 22  | X        | Y        |
| 22  | X        | Y        |

# Finding all type 1 TAD boundaries

## Here I will sort by chromosome and start site the count the duplicates across chromosome and start site

```{r echo=FALSE}
# The data is already sorted by chromosome and start site
dupstart <- arrowhead_data[duplicated(arrowhead_data[,1:2]),]
head(dupstart)
dim(dupstart)
#303   3
table(dupstart$Chromosome)
```

### There are some situations where there are multiple nested TADs

```{r echo=FALSE}
dupstart[duplicated(dupstart$Start),]
#example
#     Chromosome     Start       End
#5968       chr3  51425000  51535000

```

### For example chromosome 3 has two TADs nested inside an interval from 51425000 to 51880000
```{r echo=FALSE}
arrowhead_data[which(arrowhead_data$Start==51425000 & arrowhead_data$Chromosome=="chr3"),]

table(dupstart[duplicated(dupstart$Start),][,1])
```

# Finding all type 2 TAD boundaries

## Here I will sort by chromosome and end site the count the duplicates across chromosome and end site

```{r echo=FALSE}
dupend <- arrowhead_data[order(arrowhead_data$Chromosome,arrowhead_data$End),]
head(dupend)

dupend <- dupend[duplicated(dupend[,c(1,3)]),]
head(dupend)
dim(dupend)
#304   3
table(dupend$Chromosome)

#finding nested TADs
dupend[duplicated(dupend$end),]
#There were were no multiple nested tads here

```

# Finding all type 3 TAD boundaries

## Here I will lag the end site so as to line up starting and ending sites then search for where start=end

```{r echo=FALSE}
# creating a function to lag a column of a data frame
#lagpad <- function(x, k) {
#    if (!is.vector(x)) 
#        stop('x must be a vector')
#    if (!is.numeric(x)) 
#        stop('x must be numeric')
#    if (!is.numeric(k))
#        stop('k must be numeric')
#    if (1 != length(k))
#        stop('k must be a single number')
#    c(rep(NA, k), x)[1 : length(x)] 
#}

lagtad <- arrowhead_data %>% mutate(End=lag(End))

#setting positions where chromosomes change equal to NA
head(lagtad[ diff(c(0,lagtad$Chromosome)) != 0, ])
lagtad[ diff(c(0,lagtad$Chromosome)) != 0, ][,3] <- NA
head(lagtad[ diff(c(0,lagtad$Chromosome)) != 0, ])

lagtaddup <- lagtad[which(lagtad$Start==lagtad$End),]
head(lagtaddup)
dim(lagtaddup)
#477 3
table(lagtaddup$Chromosome)

arrowhead_data[which((arrowhead_data$Start==3540000 | arrowhead_data$End==3540000) & arrowhead_data$Chromosome=="chr1"),]
#    Chromosome   Start     End
#477       chr1 3395000 3540000
#784       chr1 3540000 3660000
```

# Finding all type 4 TAD boundaries

## Here I will simply search for duplicates across all columns

```{r echo=FALSE}
arrowhead_data[duplicated(arrowhead_data),]
#there are no duplicate intervals
```





