---
title: "Applying Mourad Model k562"
author: "Spiro Stilianoudakis"
date: "July 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Using BorderAnalysisFun function from Mourad paper to run MLR

#Loading packages
```{r}
library(pROC)
library(BSgenome.Hsapiens.UCSC.hg19)
library(HiCfeat)
```


#setting directory
```{r}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/data")
```


#The BorderAnalysisFun function has 4 key parameters:
#1. A list of GRanges objects for each annotation
#2. A character vector listing the annotation variable names
#3. A GRanges object of the TAD data
#4. A Seqinfo object that provides information regarding the genome being used


# Coordinate Data

#1 Creating a list of GRanges for each annotation

## There are no 3D subcompartments for k562 cell line



##DGV
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/DGV/DGV_files")

temp = list.files()

complex <- read.table(temp[1],header=FALSE,sep="\t",stringsAsFactors = FALSE)
complex <- complex[-which(complex$V1=="chrX" | complex$V1=="chrY" | complex$V1=="chr9"),]
deletion <- read.table(temp[2],header=FALSE,sep="\t",stringsAsFactors = FALSE)
deletion <- deletion[-which(deletion$V1=="chrX" | deletion$V1=="chrY" | deletion$V1=="chr9"),]
duplication <- read.table(temp[3],header=FALSE,sep="\t",stringsAsFactors = FALSE)
duplication <- duplication[-which(duplication$V1=="chrX" | duplication$V1=="chrY" | duplication$V1=="chr9"),]
gain_loss <- read.table(temp[4],header=FALSE,sep="\t",stringsAsFactors = FALSE)
gain_loss <- gain_loss[-which(gain_loss$V1=="chrX" | gain_loss$V1=="chrY" | gain_loss$V1=="chr9"),]
insertion <- read.table(temp[5],header=FALSE,sep="\t",stringsAsFactors = FALSE)
insertion <- insertion[-which(insertion$V1=="chrX" | insertion$V1=="chrY" | insertion$V1=="chr9"),]
inversion <- read.table(temp[6],header=FALSE,sep="\t",stringsAsFactors = FALSE)
inversion <- inversion[-which(inversion$V1=="chrX" | inversion$V1=="chrY" | inversion$V1=="chr9"),]
mobile_element_insertion <- read.table(temp[7],header=FALSE,sep="\t",stringsAsFactors = FALSE)
mobile_element_insertion <- mobile_element_insertion[-which(mobile_element_insertion$V1=="chrX" | mobile_element_insertion$V1=="chrY" | mobile_element_insertion$V1=="chr9"),]
novel_sequence_insertion <- read.table(temp[8],header=FALSE,sep="\t",stringsAsFactors = FALSE)
novel_sequence_insertion <- novel_sequence_insertion[-which(novel_sequence_insertion$V1=="chrX" | novel_sequence_insertion$V1=="chrY" | novel_sequence_insertion$V1=="chr9"),]
sequence_alteration <- read.table(temp[9],header=FALSE,sep="\t",stringsAsFactors = FALSE)
sequence_alteration <- sequence_alteration[-which(sequence_alteration$V1=="chrX" | sequence_alteration$V1=="chrY" | sequence_alteration$V1=="chr9"),]
tandem_duplication <- read.table(temp[10],header=FALSE,sep="\t",stringsAsFactors = FALSE)
tandem_duplication <- tandem_duplication[-which(tandem_duplication$V1=="chrX" | tandem_duplication$V1=="chrY" | tandem_duplication$V1=="chr9"),]

complex_gr <- GRanges(seqnames=complex$V1,IRanges(start=complex$V2,end=complex$V3))
deletion_gr <- GRanges(seqnames=deletion$V1,IRanges(start=deletion$V2,end=deletion$V3))
duplication_gr <- GRanges(seqnames=duplication$V1,IRanges(start=duplication$V2,end=duplication$V3))
gain_loss_gr <- GRanges(seqnames=gain_loss$V1,IRanges(start=gain_loss$V2,end=gain_loss$V3))
insertion_gr <- GRanges(seqnames=insertion$V1,IRanges(start=insertion$V2,end=insertion$V3))
inversion_gr <- GRanges(seqnames=inversion$V1,IRanges(start=inversion$V2,end=inversion$V3))
mobile_element_insertion_gr <- GRanges(seqnames=mobile_element_insertion$V1,IRanges(start=mobile_element_insertion$V2,end=mobile_element_insertion$V3))
novel_sequence_insertion_gr <- GRanges(seqnames=novel_sequence_insertion$V1,IRanges(start=novel_sequence_insertion$V2,end=novel_sequence_insertion$V3))
sequence_alteration_gr <- GRanges(seqnames=sequence_alteration$V1,IRanges(start=sequence_alteration$V2,end=sequence_alteration$V3))
tandem_duplication_gr <- GRanges(seqnames=tandem_duplication$V1,IRanges(start=tandem_duplication$V2,end=tandem_duplication$V3))

 

```

## GERP

```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/GERP/GERP_hg19.bed")

gerp <- read.table("GERP_hg19.BED",header=FALSE,sep="\t",stringsAsFactors = FALSE)
gerp <- gerp[-which(gerp$V1=="chrX" | gerp$V1=="chrY" | gerp$V1=="chr9"),]

gerp_gr <- GRanges(seqnames=gerp$V1,IRanges(start=gerp$V2,end=gerp$V3))


```

## nestedRepeats

```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/nestedRepeats/nestedRepeats_files")

temp = list.files()

DNA <- read.table(temp[1],header=FALSE,sep="\t",stringsAsFactors = FALSE)
DNA <- DNA[-which(DNA$V1=="chrX" | DNA$V1=="chrY" | DNA$V1=="chr9"),]
line <- read.table(temp[2],header=FALSE,sep="\t",stringsAsFactors = FALSE)
line <- line[-which(line$V1=="chrX" | line$V1=="chrY" | line$V1=="chr9"),]
low_complexity <- read.table(temp[3],header=FALSE,sep="\t",stringsAsFactors = FALSE)
low_complexity <- low_complexity[-which(low_complexity$V1=="chrX" | low_complexity$V1=="chrY" | low_complexity$V1=="chr9"),]
LTR <- read.table(temp[4],header=FALSE,sep="\t",stringsAsFactors = FALSE)
LTR <- LTR[-which(LTR$V1=="chrX" | LTR$V1=="chrY" | LTR$V1=="chr9"),]
other <- read.table(temp[5],header=FALSE,sep="\t",stringsAsFactors = FALSE)
other <- other[-which(other$V1=="chrX" | other$V1=="chrY" | other$V1=="chr9"),]
RC <- read.table(temp[6],header=FALSE,sep="\t",stringsAsFactors = FALSE)
RC <- RC[-which(RC$V1=="chrX" | RC$V1=="chrY" | RC$V1=="chr9"),]
RNA <- read.table(temp[7],header=FALSE,sep="\t",stringsAsFactors = FALSE)
RNA <- RNA[-which(RNA$V1=="chrX" | RNA$V1=="chrY" | RNA$V1=="chr9"),]
rRNA <- read.table(temp[8],header=FALSE,sep="\t",stringsAsFactors = FALSE)
rRNA <- rRNA[-which(rRNA$V1=="chrX" | rRNA$V1=="chrY" | rRNA$V1=="chr9"),]
satellite <- read.table(temp[9],header=FALSE,sep="\t",stringsAsFactors = FALSE)
satellite <- satellite[-which(satellite$V1=="chrX" | satellite$V1=="chrY" | satellite$V1=="chr9"),]
scRNA <- read.table(temp[10],header=FALSE,sep="\t",stringsAsFactors = FALSE)
scRNA <- scRNA[-which(scRNA$V1=="chrX" | scRNA$V1=="chrY" | scRNA$V1=="chr9"),]
simple_repeat <- read.table(temp[11],header=FALSE,sep="\t",stringsAsFactors = FALSE)
simple_repeat <- simple_repeat[-which(simple_repeat$V1=="chrX" | simple_repeat$V1=="chrY" | simple_repeat$V1=="chr9"),]
SINE <- read.table(temp[12],header=FALSE,sep="\t",stringsAsFactors = FALSE)
SINE <- SINE[-which(SINE$V1=="chrX" | SINE$V1=="chrY" | SINE$V1=="chr9"),]
snRNA <- read.table(temp[13],header=FALSE,sep="\t",stringsAsFactors = FALSE)
snRNA <- snRNA[-which(snRNA$V1=="chrX" | snRNA$V1=="chrY" | snRNA$V1=="chr9"),]
srpRNA <- read.table(temp[14],header=FALSE,sep="\t",stringsAsFactors = FALSE)
srpRNA <- srpRNA[-which(srpRNA$V1=="chrX" | srpRNA$V1=="chrY" | srpRNA$V1=="chr9"),]
tRNA <- read.table(temp[15],header=FALSE,sep="\t",stringsAsFactors = FALSE)
tRNA <- tRNA[-which(tRNA$V1=="chrX" | tRNA$V1=="chrY" | tRNA$V1=="chr9"),]
unknown <- read.table(temp[16],header=FALSE,sep="\t",stringsAsFactors = FALSE)
unknown <- unknown[-which(unknown$V1=="chrX" | unknown$V1=="chrY" | unknown$V1=="chr9"),]

DNA_gr <- GRanges(seqnames=DNA$V1,IRanges(start=DNA$V2,end=DNA$V3))
line_gr <- GRanges(seqnames=line$V1,IRanges(start=line$V2,end=line$V3))
low_complexity_gr <- GRanges(seqnames=low_complexity$V1,IRanges(start=low_complexity$V2,end=low_complexity$V3))
LTR_gr <- GRanges(seqnames=LTR$V1,IRanges(start=LTR$V2,end=LTR$V3))
other_gr <- GRanges(seqnames=other$V1,IRanges(start=other$V2,end=other$V3))
RC_gr <- GRanges(seqnames=RC$V1,IRanges(start=RC$V2,end=RC$V3))
RNA_gr <- GRanges(seqnames=RNA$V1,IRanges(start=RNA$V2,end=RNA$V3))
rRNA_gr <- GRanges(seqnames=rRNA$V1,IRanges(start=rRNA$V2,end=rRNA$V3))
satellite_gr <- GRanges(seqnames=satellite$V1,IRanges(start=satellite$V2,end=satellite$V3))
scRNA_gr <- GRanges(seqnames=scRNA$V1,IRanges(start=scRNA$V2,end=scRNA$V3))
simple_repeat_gr <- GRanges(seqnames=simple_repeat$V1,IRanges(start=simple_repeat$V2,end=simple_repeat$V3))
SINE_gr <- GRanges(seqnames=SINE$V1,IRanges(start=SINE$V2,end=SINE$V3))
snRNA_gr <- GRanges(seqnames=snRNA$V1,IRanges(start=snRNA$V2,end=snRNA$V3))
srpRNA_gr <- GRanges(seqnames=srpRNA$V1,IRanges(start=srpRNA$V2,end=srpRNA$V3))
tRNA_gr <- GRanges(seqnames=tRNA$V1,IRanges(start=tRNA$V2,end=tRNA$V3))
unknown_gr <- GRanges(seqnames=unknown$V1,IRanges(start=unknown$V2,end=unknown$V3))

```


## super_enhancers
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/super_enhancers")

temp = list.files()

se_k562 <- read.table(temp[1],header=FALSE,sep="\t",stringsAsFactors = FALSE)
se_k562 <- se_k562[-which(se_k562$V1=="chrX" | se_k562$V1=="chrY" | se_k562$V1=="chr9"),]

se_k562_gr <- GRanges(seqnames=se_k562$V1,IRanges(start=se_k562$V2,end=se_k562$V3))

 
```

## UCNE
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/UCNEs/UCNE.bed")

UCNE <- read.table("UCNE.BED",header=FALSE,sep="\t",stringsAsFactors = FALSE)
UCNE <- UCNE[-which(UCNE$V1=="chrX" | UCNE$V1=="chrY" | UCNE$V1=="chr9"),]

UCNE_gr <- GRanges(seqnames=UCNE$V1,IRanges(start=UCNE$V2,end=UCNE$V3))


```

## VMR
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/VMRs")

VMR <- read.table("VMR_hg19.BED",header=FALSE,sep="\t",stringsAsFactors = FALSE)
VMR <- VMR[-which(VMR$V1=="chrX" | VMR$V1=="chrY" | VMR$V1=="chr9"),]

VMR_gr <- GRanges(seqnames=VMR$V1,IRanges(start=VMR$V2,end=VMR$V3))

```

## BroadHMM
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/BroadHmm")

temp <- list.files()

#k562
k562_TxnElongation <- read.table(temp[2],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_TxnElongation <- k562_TxnElongation[-which(k562_TxnElongation$V1=="chrX" | k562_TxnElongation$V1=="chrY" | k562_TxnElongation$V1=="chr9"),]
k562_WeakTxn <- read.table(temp[3],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_WeakTxn <- k562_WeakTxn[-which(k562_WeakTxn$V1=="chrX" | k562_WeakTxn$V1=="chrY" | k562_WeakTxn$V1=="chr9"),]
k562_Repressed <- read.table(temp[4],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_Repressed <- k562_Repressed[-which(k562_Repressed$V1=="chrX" | k562_Repressed$V1=="chrY" | k562_Repressed$V1=="chr9"),]
k562_Heterochromlo <- read.table(temp[5],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_Heterochromlo <- k562_Heterochromlo[-which(k562_Heterochromlo$V1=="chrX" | k562_Heterochromlo$V1=="chrY" | k562_Heterochromlo$V1=="chr9"),]
k562_RepetitiveCNV14 <- read.table(temp[6],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_RepetitiveCNV14 <- k562_RepetitiveCNV14[-which(k562_RepetitiveCNV14$V1=="chrX" | k562_RepetitiveCNV14$V1=="chrY" | k562_RepetitiveCNV14$V1=="chr9"),]
k562_RepetitiveCNV15 <- read.table(temp[7],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_RepetitiveCNV15 <- k562_RepetitiveCNV15[-which(k562_RepetitiveCNV15$V1=="chrX" | k562_RepetitiveCNV15$V1=="chrY" | k562_RepetitiveCNV15$V1=="chr9"),]
k562_ActivePromoter <- read.table(temp[8],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_ActivePromoter <- k562_ActivePromoter[-which(k562_ActivePromoter$V1=="chrX" | k562_ActivePromoter$V1=="chrY" | k562_ActivePromoter$V1=="chr9"),]
k562_WeakPromoter <- read.table(temp[9],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_WeakPromoter <- k562_WeakPromoter[-which(k562_WeakPromoter$V1=="chrX" | k562_WeakPromoter$V1=="chrY" | k562_WeakPromoter$V1=="chr9"),]
k562_PoisedPromoter <- read.table(temp[10],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_PoisedPromoter <- k562_PoisedPromoter[-which(k562_PoisedPromoter$V1=="chrX" | k562_PoisedPromoter$V1=="chrY" | k562_PoisedPromoter$V1=="chr9"),]
k562_StrongEnhancer4 <- read.table(temp[11],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_StrongEnhancer4 <- k562_StrongEnhancer4[-which(k562_StrongEnhancer4$V1=="chrX" | k562_StrongEnhancer4$V1=="chrY" | k562_StrongEnhancer4$V1=="chr9"),]
k562_StrongEnhancer5 <- read.table(temp[12],header=FALSE,sep="\t",stringsAsFactors = FALSE) 
k562_StrongEnhancer5 <- k562_StrongEnhancer5[-which(k562_StrongEnhancer5$V1=="chrX" | k562_StrongEnhancer5$V1=="chrY" | k562_StrongEnhancer5$V1=="chr9"),]
k562_WeakEnhancer6 <- read.table(temp[13],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_WeakEnhancer6 <- k562_WeakEnhancer6[-which(k562_WeakEnhancer6$V1=="chrX" | k562_WeakEnhancer6$V1=="chrY" | k562_WeakEnhancer6$V1=="chr9"),]
k562_WeakEnhancer7 <- read.table(temp[14],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_WeakEnhancer7 <- k562_WeakEnhancer7[-which(k562_WeakEnhancer7$V1=="chrX" | k562_WeakEnhancer7$V1=="chrY" | k562_WeakEnhancer7$V1=="chr9"),]
k562_Insulator <- read.table(temp[15],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_Insulator <- k562_Insulator[-which(k562_Insulator$V1=="chrX" | k562_Insulator$V1=="chrY" | k562_Insulator$V1=="chr9"),]
k562_TxnTransition <- read.table(temp[16],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_TxnTransition <- k562_TxnTransition[-which(k562_TxnTransition$V1=="chrX" | k562_TxnTransition$V1=="chrY" | k562_TxnTransition$V1=="chr9"),]

k562_TxnElongation_gr <- GRanges(seqnames=k562_TxnElongation$V1,IRanges(start=k562_TxnElongation$V2,end=k562_TxnElongation$V3))
k562_WeakTxn_gr <- GRanges(seqnames=k562_WeakTxn$V1,IRanges(start=k562_WeakTxn$V2,end=k562_WeakTxn$V3))
k562_Repressed_gr <- GRanges(seqnames=k562_Repressed$V1,IRanges(start=k562_Repressed$V2,end=k562_Repressed$V3))
k562_Heterochromlo_gr <- GRanges(seqnames=k562_Heterochromlo$V1,IRanges(start=k562_Heterochromlo$V2,end=k562_Heterochromlo$V3)) 
k562_RepetitiveCNV14_gr <- GRanges(seqnames=k562_RepetitiveCNV14$V1,IRanges(start=k562_RepetitiveCNV14$V2,end=k562_RepetitiveCNV14$V3)) 
k562_RepetitiveCNV15_gr <- GRanges(seqnames=k562_RepetitiveCNV15$V1,IRanges(start=k562_RepetitiveCNV15$V2,end=k562_RepetitiveCNV15$V3))
k562_ActivePromoter_gr <- GRanges(seqnames=k562_ActivePromoter$V1,IRanges(start=k562_ActivePromoter$V2,end=k562_ActivePromoter$V3))
k562_WeakPromoter_gr <- GRanges(seqnames=k562_WeakPromoter$V1,IRanges(start=k562_WeakPromoter$V2,end=k562_WeakPromoter$V3))
k562_PoisedPromoter_gr <- GRanges(seqnames=k562_PoisedPromoter$V1,IRanges(start=k562_PoisedPromoter$V2,end=k562_PoisedPromoter$V3)) 
k562_StrongEnhancer4_gr <- GRanges(seqnames=k562_StrongEnhancer4$V1,IRanges(start=k562_StrongEnhancer4$V2,end=k562_StrongEnhancer4$V3)) 
k562_StrongEnhancer5_gr <- GRanges(seqnames=k562_StrongEnhancer5$V1,IRanges(start=k562_StrongEnhancer5$V2,end=k562_StrongEnhancer5$V3))
k562_WeakEnhancer6_gr <- GRanges(seqnames=k562_WeakEnhancer6$V1,IRanges(start=k562_WeakEnhancer6$V2,end=k562_WeakEnhancer6$V3)) 
k562_WeakEnhancer7_gr <- GRanges(seqnames=k562_WeakEnhancer7$V1,IRanges(start=k562_WeakEnhancer7$V2,end=k562_WeakEnhancer7$V3)) 
k562_Insulator_gr <- GRanges(seqnames=k562_Insulator$V1,IRanges(start=k562_Insulator$V2,end=k562_Insulator$V3))
k562_TxnTransition_gr <- GRanges(seqnames=k562_TxnTransition$V1,IRanges(start=k562_TxnTransition$V2,end=k562_TxnTransition$V3)) 

  
```

## Combined
```{r include=FALSE}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/Combined")

temp <- list.files()

#k562
k562_CTCF <- read.table(temp[2],header=FALSE,sep="\t",stringsAsFactors = FALSE) 
k562_CTCF <- k562_CTCF[-which(k562_CTCF$V1=="chrX" | k562_CTCF$V1=="chrY" | k562_CTCF$V1=="chr9"),]
k562_E <- read.table(temp[3],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_E <- k562_E[-which(k562_E$V1=="chrX" | k562_E$V1=="chrY" | k562_E$V1=="chr9"),]
k562_PF <- read.table(temp[4],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_PF <- k562_PF[-which(k562_PF$V1=="chrX" | k562_PF$V1=="chrY" | k562_PF$V1=="chr9"),]
k562_R <- read.table(temp[5],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_R <- k562_R[-which(k562_R$V1=="chrX" | k562_R$V1=="chrY" | k562_R$V1=="chr9"),]
k562_T <- read.table(temp[6],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_T <- k562_T[-which(k562_T$V1=="chrX" | k562_T$V1=="chrY" | k562_T$V1=="chr9"),]
k562_TSS <- read.table(temp[7],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_TSS <- k562_TSS[-which(k562_TSS$V1=="chrX" | k562_TSS$V1=="chrY" | k562_TSS$V1=="chr9"),]
k562_WE <- read.table(temp[8],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_WE <- k562_WE[-which(k562_WE$V1=="chrX" | k562_WE$V1=="chrY" | k562_WE$V1=="chr9"),]

k562_CTCF_gr <- GRanges(seqnames=k562_CTCF$V1,IRanges(start=k562_CTCF$V2,end=k562_CTCF$V3))
k562_E_gr <- GRanges(seqnames=k562_E$V1,IRanges(start=k562_E$V2,end=k562_E$V3))
k562_PF_gr <- GRanges(seqnames=k562_PF$V1,IRanges(start=k562_PF$V2,end=k562_PF$V3))
k562_R_gr <- GRanges(seqnames=k562_R$V1,IRanges(start=k562_R$V2,end=k562_R$V3))
k562_T_gr <- GRanges(seqnames=k562_T$V1,IRanges(start=k562_T$V2,end=k562_T$V3))
k562_TSS_gr <- GRanges(seqnames=k562_TSS$V1,IRanges(start=k562_TSS$V2,end=k562_TSS$V3))
k562_WE_gr <- GRanges(seqnames=k562_WE$V1,IRanges(start=k562_WE$V2,end=k562_WE$V3))

```

## DNase I

```{r}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/DNaseI")

temp <- list.files()

k562_DNaseI <- read.table(temp[1],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_DNaseI <- k562_DNaseI[-which(k562_DNaseI$V1=="chrX" | k562_DNaseI$V1=="chrY" | k562_DNaseI$V1=="chr9"),]

k562_DNaseI_gr <- GRanges(seqnames=k562_DNaseI$V1,IRanges(start=k562_DNaseI$V2,end=k562_DNaseI$V3))


```

## Histone Modifications

```{r}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/annotations/HistoneModifications")

temp <- list.files()

k562_H2az <- read.table(temp[1],header=FALSE,sep="\t",stringsAsFactors = FALSE) 
k562_H2az <- k562_H2az[-which(k562_H2az$V1=="chrX" | k562_H2az$V1=="chrY" | k562_H2az$V1=="chr9"),]
k562_H3k27ac <- read.table(temp[2],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k27ac <- k562_H3k27ac[-which(k562_H3k27ac$V1=="chrX" | k562_H3k27ac$V1=="chrY" | k562_H3k27ac$V1=="chr9"),]
k562_H3k27me3 <- read.table(temp[3],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k27me3 <- k562_H3k27me3[-which(k562_H3k27me3$V1=="chrX" | k562_H3k27me3$V1=="chrY" | k562_H3k27me3$V1=="chr9"),]
k562_H3k36me3 <- read.table(temp[4],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k36me3 <- k562_H3k36me3[-which(k562_H3k36me3$V1=="chrX" | k562_H3k36me3$V1=="chrY" | k562_H3k36me3$V1=="chr9"),]
k562_H3k4me1 <- read.table(temp[5],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k4me1 <- k562_H3k4me1[-which(k562_H3k4me1$V1=="chrX" | k562_H3k4me1$V1=="chrY" | k562_H3k4me1$V1=="chr9"),] 
k562_H3k4me2 <- read.table(temp[6],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k4me2 <- k562_H3k4me2[-which(k562_H3k4me2$V1=="chrX" | k562_H3k4me2$V1=="chrY" | k562_H3k4me2$V1=="chr9"),]
k562_H3k4me3 <- read.table(temp[7],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k4me3 <- k562_H3k4me3[-which(k562_H3k4me3$V1=="chrX" | k562_H3k4me3$V1=="chrY" | k562_H3k4me3$V1=="chr9"),]
k562_H3k79me2 <- read.table(temp[8],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k79me2 <- k562_H3k79me2[-which(k562_H3k79me2$V1=="chrX" | k562_H3k79me2$V1=="chrY" | k562_H3k79me2$V1=="chr9"),]
k562_H3k9ac <- read.table(temp[9],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k9ac <- k562_H3k9ac[-which(k562_H3k9ac$V1=="chrX" | k562_H3k9ac$V1=="chrY" | k562_H3k9ac$V1=="chr9"),]
k562_H3k9me3 <- read.table(temp[10],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H3k9me3 <- k562_H3k9me3[-which(k562_H3k9me3$V1=="chrX" | k562_H3k9me3$V1=="chrY" | k562_H3k9me3$V1=="chr9"),]
k562_H4k20me1 <- read.table(temp[11],header=FALSE,sep="\t",stringsAsFactors = FALSE)
k562_H4k20me1 <- k562_H4k20me1[-which(k562_H4k20me1$V1=="chrX" | k562_H4k20me1$V1=="chrY" | k562_H4k20me1$V1=="chr9"),]

k562_H2az_gr <- GRanges(seqnames=k562_H2az$V1,IRanges(start=k562_H2az$V2,end=k562_H2az$V3))
k562_H3k27ac_gr <- GRanges(seqnames=k562_H3k27ac$V1,IRanges(start=k562_H3k27ac$V2,end=k562_H3k27ac$V3))
k562_H3k27me3_gr <- GRanges(seqnames=k562_H3k27me3$V1,IRanges(start=k562_H3k27me3$V2,end=k562_H3k27me3$V3))
k562_H3k36me3_gr <- GRanges(seqnames=k562_H3k36me3$V1,IRanges(start=k562_H3k36me3$V2,end=k562_H3k36me3$V3))
k562_H3k4me1_gr <- GRanges(seqnames=k562_H3k4me1$V1,IRanges(start=k562_H3k4me1$V2,end=k562_H3k4me1$V3))
k562_H3k4me2_gr <- GRanges(seqnames=k562_H3k4me2$V1,IRanges(start=k562_H3k4me2$V2,end=k562_H3k4me2$V3))
k562_H3k4me3_gr <- GRanges(seqnames=k562_H3k4me3$V1,IRanges(start=k562_H3k4me3$V2,end=k562_H3k4me3$V3))
k562_H3k79me2_gr <- GRanges(seqnames=k562_H3k79me2$V1,IRanges(start=k562_H3k79me2$V2,end=k562_H3k79me2$V3))
k562_H3k9ac_gr <- GRanges(seqnames=k562_H3k9ac$V1,IRanges(start=k562_H3k9ac$V2,end=k562_H3k9ac$V3))
k562_H3k9me3_gr <- GRanges(seqnames=k562_H3k9me3$V1,IRanges(start=k562_H3k9me3$V2,end=k562_H3k9me3$V3))
k562_H4k20me1_gr <- GRanges(seqnames=k562_H4k20me1$V1,IRanges(start=k562_H4k20me1$V2,end=k562_H4k20me1$V3))

```


## Concatinating all GRanges into a list
```{r}
genomicFeatureList.GR <- list(complex=complex_gr, 
deletion=deletion_gr, 
duplication=duplication_gr, 
gain_loss=gain_loss_gr, 
insertion=insertion_gr, 
inversion=inversion_gr, 
mobile_element_insertion=mobile_element_insertion_gr, 
novel_sequence_insertion=novel_sequence_insertion_gr, 
sequence_alteration=sequence_alteration_gr, 
tandem_duplication=tandem_duplication_gr, 
gerp=gerp_gr, 
DNA=DNA_gr, 
line=line_gr, 
low_complexity=low_complexity_gr, 
LTR=LTR_gr, 
other=other_gr, 
RC=RC_gr, 
RNA=RNA_gr, 
rRNA=rRNA_gr, 
satellite=satellite_gr, 
scRNA=scRNA_gr, 
simple_repeat=simple_repeat_gr, 
SINE=SINE_gr, 
snRNA=snRNA_gr, 
srpRNA=srpRNA_gr, 
tRNA=tRNA_gr, 
unknown=unknown_gr, 
se=se_k562_gr,  
UCNE=UCNE_gr, 
VMR=VMR_gr,
TxnElongation=k562_TxnElongation_gr, 
WeakTxn=k562_WeakTxn_gr, 
Repressed=k562_Repressed_gr, 
Heterochromlo=k562_Heterochromlo_gr,  
RepetitiveCNV14=k562_RepetitiveCNV14_gr, 
RepetitiveCNV15=k562_RepetitiveCNV15_gr, 
ActivePromoter=k562_ActivePromoter_gr, 
WeakPromoter=k562_WeakPromoter_gr, 
PoisedPromoter=k562_PoisedPromoter_gr, 
StrongEnhancer4=k562_StrongEnhancer4_gr,  
StrongEnhancer5=k562_StrongEnhancer5_gr, 
WeakEnhancer6=k562_WeakEnhancer6_gr, 
WeakEnhancer7=k562_WeakEnhancer7_gr, 
Insulator=k562_Insulator_gr, 
TxnTransition=k562_TxnTransition_gr, 
CTCF=k562_CTCF_gr, 
E=k562_E_gr, 
PF=k562_PF_gr, 
R=k562_R_gr, 
T=k562_T_gr, 
TSS=k562_TSS_gr, 
WE=k562_WE_gr, 
DNaseI=k562_DNaseI_gr, 
H2az=k562_H2az_gr, 
H3k27ac=k562_H3k27ac_gr, 
H3k27me3=k562_H3k27me3_gr, 
H3k36me3=k562_H3k36me3_gr, 
H3k4me1=k562_H3k4me1_gr,
H3k4me2=k562_H3k4me2_gr, 
H3k4me3=k562_H3k4me3_gr,
H3k79me2=k562_H3k79me2_gr, 
H3k9ac=k562_H3k9ac_gr, 
H3k9me3=k562_H3k9me3_gr, 
H4k20me1=k562_H4k20me1_gr) 

#removing annotations with less than 22 chromosomes
for(i in 1:length(names(genomicFeatureList.GR))){
   print(length(table(seqnames(genomicFeatureList.GR[[i]])))) 
 }
genomicFeatureList.GR <- genomicFeatureList.GR[-c(18,19,21,24,26)]

#removing chromosome X and Y
#chrs <- paste("chr",c(1,10:19,2,20:22,3:9),sep = "")
#for(i in 1:length(names(genomicFeatureList.GR))){
#  if(length(table(seqnames(genomicFeatureList.GR[[i]]))) > 22){
#    genomicFeatureList.GR[[i]] <- genomicFeatureList.GR[[i]][which(as.character(seqnames(genomicFeatureList.GR[[i]])) %in% chrs)]
    #levels(seqnames(genomicFeatureList.GR[[i]])) <- chrs
#  }
#}



```

#2. A character vector listing the annotation variable names
```{r}
annotNames <- names(genomicFeatureList.GR)

```

#3. A GRanges object of the TAD data
```{r}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/data")

#Reading in TAD data for k562 cell line
tad <- read.table("GSE63525_K562_Arrowhead_domainlist.txt",header = FALSE, stringsAsFactors = FALSE, skip = 1)
tad$V1 <- paste("chr",tad$V1,sep = "")
tad$V2 <- as.numeric(tad$V2)
tad$V3 <- as.numeric(tad$V3)
#removing X/Y chromosome
tad <- tad[-which(tad$V1=="chrX" | tad$V1=="chrY"),]
#Formating the data
tad <- tad[order(tad$V1,tad$V2),]


domains.GR <- GRanges(seqnames=tad$V1,IRanges(start=tad$V2,end=tad$V3))


```

#4. A Seqinfo object that provides information regarding the genome being used
```{r}

seqn <- unique(tad$V1)

genome <- BSgenome.Hsapiens.UCSC.hg19
#there is no 9 chr in tad data for k562
seql <- seqlengths(genome)[c(1:8,10:22)]
seql <- seql[c(1,9:18,2,19:21,3:8)]
seql <- as.vector(seql)
isCirc <- isCircular(genome)[c(1,9:18,2,19:21,3:8)]
isCirc <- as.vector(isCirc)
seqInfoChr <- Seqinfo(seqn, seqlengths=seql, 
                      isCircular=isCirc, 
                      genome="hg19")

#providing seqlengths to each grange object in the list
for(i in 1:length(annotNames)){
  seqlengths(genomicFeatureList.GR[[i]]) <- seqlengths(seqInfoChr)
}


```

# Saving objects

```{r}
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/k562")

saveRDS(genomicFeatureList.GR, "genomicFeatureList.GR.k562.rds")
saveRDS(annotNames, "annotNames.k562.rds")
saveRDS(domains.GR, "domains.GR.k562.rds")
saveRDS(seqInfoChr, "seqInfoChr.k562.rds")
```




BA_res=borderAnalysisFun(genomicFeatureList.GR=genomicFeatureList.GR,
                         GFDataType="bed",
                         annotNames=annotNames,
                         domains.GR=domains.GR,
                         seqInfoChr=seqInfoChr,
                         analysisMode="MLR",
                         binSize=1000,borderSize=1000,
                         LRT=FALSE,
                         interactionTerms="",
                         testBorderType=FALSE,
                         verbose=FALSE)


prob=predict(Analysisi,type=c("response"))
binMat.mati$prob=prob
g <- roc(Border ~ prob, data = binMat.mati)
pROC::auc(g)

simple_roc <- function(labels, scores){
        labels <- labels[order(scores, decreasing=TRUE)]
        data.frame(TPR=cumsum(labels)/sum(labels), FPR=cumsum(!labels)/sum(!labels), labels)
      }
prob=predict(Analysisi,type=c("response"))
head(prob)
roc <- simple_roc(binMat.mati$Border, binMat.mati$prob)
head(roc)
plot(roc$FPR,roc$TPR, type="l")
abline(0,1)

library(zoo)
library(pracma)

trapz(roc$FPR,roc$TPR)
#0.7493966
sum(diff(roc$FPR)*rollmean(roc$TPR,2))
#0.7493966
