---
title: "TAD Boundary Analysis V2"
author: "Spiro Stilianoudakis"
date: "April 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading Libraries

```{r}
#library(MultiAssayExperiment)
library(GenomicRanges)
#library(IRanges)
library(caret)
library(data.table)
library(gbm)
library(pROC)
library(plyr)
library(dplyr)
```


# Reading in TAD data

```{r}

#setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/data")
setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data_analysis/data")

domains <- read.table("arrowhead_data.txt", header=T)
domains <- domains[,1:3]
head(domains)
dim(domains)
#9274   3

```

# Creating TAD boundaries

## The TAD boundary start and end coordinates were row binded into one long vector. The vector was sorted and each coordinate was flanked on each side by 500 bases for a 1kb TAD boundary bin.

```{r}

#Removing the X chromosome from the domain data
domains <- domains[-which(domains$Chromosome=="chrX"),]

#Sorting the numeric chromosome coordinates
domains <- domains[order(as.numeric(substr(domains$Chromosome,4,5)), domains$Start),]

#adding distance variable
domains$distance <- domains$End - domains$Start
#log2 transform of distance
domains$logdist <- log(domains$distance, base = 2)

#saving domain data
saveRDS(domains, "domains.rds")

#plotting distances grouped by chromosome
ggplot() + 
  geom_density(data=domains, aes(x=distance, group=Chromosome, color=Chromosome)) + 
  xlab("Length of TADs") +
  ylab("Density")+
  theme_minimal()

ggplot() + 
  geom_density(data=domains, aes(x=logdist, group=Chromosome, color=Chromosome)) + 
  xlab("Length of TADs") +
  ylab("Density")+
  theme_minimal()

# Creating a GRanges object out of the tads

coords <- domains
colnames(coords)[2:3] <- c("coordinate", "coordinate")
coords <- rbind.data.frame(coords[,c(1,2)],coords[,c(1,3)])


#remove duplicates for coordinates that are conjoined
coords <- coords[!duplicated(coords),]
coords <- coords[order(as.numeric(substr(coords$Chromosome,4,5))),]
dim(coords)
#16770     2



# flanking either side of the TAD boundary by 500 bases for a 1kb centered boundary region
coords$Chromosome <- as.character(coords$Chromosome)
bounds <- GRanges(seqnames=coords$Chromosome, ranges=IRanges(start=coords$coordinate, width=1))
bounds <- resize(bounds, 1000, fix = "center")
bounds
prop.table(table(seqnames(bounds)))


setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878")
saveRDS(bounds, "bounds.rds")

```

# Reading in and cleaning genomic feature data (subcompartments)

```{r}

# setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/data")

# subcompart <- read.table("GSE63525_GM12878_subcompartments.BED",header = F)
# subcompart <- subcompart[,1:4]
# names(subcompart) <- c("chr", "x1", "x2", "compartment")
# dim(subcompart)
# subcompart <- subcompart[-which(is.na(subcompart$compartment)),]
# head(subcompart)
# table(subcompart$compartment)
# subcompart$compartment <- as.character(subcompart$compartment)
# 
# #combining compartments
# subcompart$compartment[grep("A",subcompart$compartment)] <- "A"
# subcompart$compartment[grep("B",subcompart$compartment)] <- "B"
# table(subcompart$compartment)
# #   A    B 
# #1739 2110
# prop.table(table(subcompart$compartment))
# #        A         B 
# #0.4518057 0.5481943 
# 
# #ordering the subcompartment data
# subcompart <- subcompart[order(substr(subcompart$chr,4,5),
#                                subcompart$x1),]
# 
# #creating granges object out of subcompartment data
# #coding strand according to A or B compartment (need to have this
# #information in order for the reduce() command to work properly) 
# #using the reduce function to collapse the consecutive intervals
# #Note: A -> +; B -> -
# subcompartgr <- GRanges(seqnames = subcompart$chr, 
#                         ranges = IRanges(start=subcompart$x1, 
#                                          end=subcompart$x2), 
#                         strand = ifelse(subcompart$compartment=="A","+","-"))
# subcompartgr <- reduce(subcompartgr)
# #reduce command shrunk the subcompartment intervals from 3849 to 2909
# #adding the type of subcompartment as metadata for granges object
# mcols(subcompartgr)$compartment <- ifelse(strand(subcompartgr)=="+","A","B")
# #getting rid of strand information
# strand(subcompartgr) <- "*"
# 
# setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData")
# 
# saveRDS(subcompartgr, "subcompartgr.rds")
# 
# #subcompartdf <- data.frame(chr = seqnames(subcompartgr),
# #                           start = start(subcompartgr),
# #                           end = end(subcompartgr),
# #                           comparment = mcols(subcompartgr)$compartment)
# #write.table(subcompartdf, "subcompart_reduced.bed")


```

# Finding where the 1kb flanked tad boundaries overlap the subcompartments and labeling the overlap as A or B

```{r}

# #There are 3 scenarios of overlaps that can occur:
# #i. partial
# #ii. within
# #iii. none
# #For partial overlaps, they can either be between two (or more) subcompartments, or with only one subcompartment (i.e. there is a break between compartments)
# 
# #For Within:
# within <- findOverlaps(bounds, subcompartgr, type="within")
# #there are 16593 intervals from the tad boundaries that overlab strictly within flanked subcompartments
# subcompoverlapwithin <- bounds[queryHits(within)]
# #attaching the subcompartment info to the within overlaps
# mcols(subcompoverlapwithin)$compartment <- mcols(subcompartgr)$compartment[subjectHits(within)]
# #adding percent overlap for each A abd B compartments
# mcols(subcompoverlapwithin)$percentA <- ifelse(mcols(subcompoverlapwithin)$compartment=="A",1,0)
# mcols(subcompoverlapwithin)$percentB <- ifelse(mcols(subcompoverlapwithin)$compartment=="B",1,0)
# 
# 
# #For none:
# subcompoverlapnone <- bounds[which(bounds %outside% subcompartgr)]
# #there are 19 intervals that have no overlaps
# #attaching the subcompartment info in the form of "N" for no overlap
# mcols(subcompoverlapnone)$compartment <- "N"
# #adding percent overlap
# mcols(subcompoverlapnone)$percentA <- 0
# mcols(subcompoverlapnone)$percentB <- 0
# 
# 
# #For partial: the difference among the sets of any type of overlap and within overlaps
# subcompoverlapany <- findOverlaps(bounds, subcompartgr, type="any")
# #there are 16904 overlaps from the tad boundaries with subcompartment intervals
# #note: there are only 16770 tad boundaries though.
# #that means that there are some places where a flanked tad boundary is a part of two (or more) subcompartment intervals
# 
# #we can fnd these by using setdiff
# partial <- setdiff(subcompoverlapany,within)
# #For example, there are 311 tad bounds that partially overlap with subcompartments
# length(unique(queryHits(partial)))
# bounds[158] #ovelaps the following two subcompartment intervals 
# subcompartgr[c(12,150)]


```

# Determing the percentage of partial overlaps

```{r}
# #determining the frequency of partial overlaps
# table(queryHits(partial))
# table(table(queryHits(partial)))
# #  1   2 
# #  5 153 
# #we see most partial overlaps are between two subcompartment intervals
# #5 out of 158 unique overlaps are with one subcompartment
# #i.e. the part of the overlap is with a particular subcompartment
# #the other part overlaps where there is no subcompartment
# 
# #investigating where it is with only 1 subcompartment interval
# which(table(queryHits(partial))==1)
# #get the location of the bounds and subcompartments that overlap
# partial[which(queryHits(partial)==5034)]
# #the subjectHits tells the postion on the subcompartments where the overlap occurs
# #example:
# bounds[5034]
# subcompartgr[2217]
# #here we have a partial overlap with only one subcompartment interval
# #since there is a break in the interval infront of subcompartgr[2216]
# 
# #we need to quantify the partial overlaps in terms of a percentage
# 
# #For overlaps between two subcompartments there is a 1-to-1 relationship between percent compartment1 (A or B) overlap with percent compartment2 (A or B) overlap. That is:
# # %C1 = 1-%C2
# 
# #For overlaps with 1 subcompartment interval the situation is different
# #Either %C1 = ? and %C2=0 
# #or
# #%C1 = 0 and %C2=?
# 
# #First consider where the overlap is with only 1 subcompartment interval
# as.numeric(rownames(as.matrix(which(table(queryHits(partial))==1))))
# oneinterval <- bounds[as.numeric(rownames(as.matrix(which(table(queryHits(partial))==1))))]
# poneinterval <- findOverlapPairs(oneinterval, subcompartgr)
# poneinterval <- pintersect(poneinterval)
# 
# overlaps1 <- findOverlaps(oneinterval, subcompartgr)
# mcols(oneinterval)$compartment <- mcols(subcompartgr)$compartment[subjectHits(overlaps1)]
# #attching compartment percent
# mcols(oneinterval)$percentA <- ifelse(mcols(oneinterval)$compartment=="A",
#                                       round(width(poneinterval)/width(oneinterval),2),
#                                       0)
# mcols(oneinterval)$percentB <- ifelse(mcols(oneinterval)$compartment=="B",
#                                       round(width(poneinterval)/width(oneinterval),2),
#                                       0)
# 
# 
# #Now we consider the overlaps between two subcompartment intervals
# twointerval <- bounds[as.numeric(rownames(as.matrix(which(table(queryHits(partial))==2))))]
# ptwointerval <- findOverlapPairs(twointerval, subcompartgr)
# ptwointerval <- pintersect(ptwointerval)
# 
# overlaps2 <- findOverlaps(twointerval, subcompartgr)
# mcols(twointerval)$compartment <- "C"
# mcols(twointerval)$percentA <- round(width(ptwointerval[!duplicated(queryHits(overlaps2))])/width(twointerval),2)
# mcols(twointerval)$percentB <- round(width(ptwointerval[!duplicated(queryHits(overlaps2))])/width(twointerval),2)
# 
# #Note: all percentage of overlaps appear to be 50%


```


# Concatinating all types of overlaps into 1 granges object

```{r}

# #within + partial(one) + partial(two) + none = 16593 + 5 + 153 + 19 = 16770
# tad_subcomp <- GRangesList(subcompoverlapwithin, subcompoverlapnone, oneinterval, twointerval)
# tad_subcomp <- unlist(tad_subcomp)
# tad_subcomp <- tad_subcomp[order(substr(seqnames(tad_subcomp),4,5), start(tad_subcomp))]
# 
# #attaching a class variable y to denote that a tad boundary exists in the flanked interval
# mcols(tad_subcomp)$y <- 1

```


# Binning the genome

```{r}
#Creating 10kb bin from min-5000 to max+5000 for each chromosome
#We first create 22 empty GRange objects to fill in with 10kb intervals
#for each of the 22 chromosomes
#This is done to create the class variable (response) Y
#Y=1 if the flanked tad boundary overlaps with a bin
#Y=0 if it is not


bins <- rep( list(GRangesList()), length(unique(coords$Chromosome)) )

for(i in 1:length(unique(coords$Chromosome))){
  seqn <- unique(coords$Chromosome)[i]

  midpt <- ((min(coords$coordinate[which(coords$Chromosome==unique(coords$Chromosome)[i])])-500) + 
            (max(coords$coordinate[which(coords$Chromosome==unique(coords$Chromosome)[i])])+500))/2

  bins[[i]] <- GRanges(seqnames=seqn, 
                ranges=IRanges(start=seq(min(coords$coordinate[which(coords$Chromosome==unique(coords$Chromosome)[i])])-500, 
                                         max(coords$coordinate[which(coords$Chromosome==unique(coords$Chromosome)[i])])+500,
                                         1000),
                               width=1000))

}

binslist <- GRangesList(bins)
binslist <- unlist(binslist)

#function to bin genome
binFunc <- function(chroms, boundpts, kb){
  #create empty list
  grlist <- rep( list(GRangesList()), length(unique(chroms)) )
  
  for(i in 1:length(unique(chroms))){
    
    #creating seqnames object for granges
    seqn <- unique(chroms)[i]
    
    #identifying minimum and maximum values to establish sequence for iranges object
    min.bound <- min(boundpts[which(chroms==unique(chroms)[i])])-(kb/2)
    max.bound <- max(boundpts[which(chroms==unique(chroms)[i])])+(kb/2)
    
    grlist[[i]] <- GRanges(seqnames=seqn,
                           ranges=IRanges(start=seq(min.bound,
                                                    max.bound,
                                                    kb),
                                          width = kb))
  }
  
  grlist <- GRangesList(grlist)
  grlist <- unlist(grlist)
  
}

#1kb bins
binslist1 <- binFunc(chroms = coords$Chromosome, boundpts = coords$coordinate, kb = 1000)
y <- countOverlaps(binslist1, bounds)
length(y) #2766314
table(y)
#       0       1 
# 2749544   16770 
prop.table(table(y))
#           0           1 
# 0.993937781 0.006062219 
mcols(binslist1)$y <- y


#5kb bins
binslist5 <- binFunc(chroms = coords$Chromosome, boundpts = coords$coordinate, kb = 5000)
y <- countOverlaps(binslist5, bounds)
length(y) #553298
table(y)
#      0      1 
# 536528  16770 
prop.table(table(y))
#          0          1 
# 0.96969084 0.03030916
mcols(binslist5)$y <- y


#10kb bins
binslist10 <- binFunc(chroms = coords$Chromosome, boundpts = coords$coordinate, kb = 10000)
y <- countOverlaps(binslist10, bounds)
length(y) #276666
table(y)
#      0      1      2      3 
# 253830  20590   2166     80 
prop.table(table(y))
#            0            1            2            3 
# 0.9174600421 0.0744218661 0.0078289345 0.0002891573 
y <- ifelse(y>=1,1,0)
prop.table(table(y))
mcols(binslist10)$y <- y


#50kb bins
binslist50 <- binFunc(chroms = coords$Chromosome, boundpts = coords$coordinate, kb = 50000)
y <- countOverlaps(binslist50, bounds)
length(y) #55361
table(y)
prop.table(table(y))
y <- ifelse(y>=1,1,0)
prop.table(table(y))
#        0         1 
# 0.7537797 0.2462203
mcols(binslist50)$y <- y


#100kb bins
binslist100 <- binFunc(chroms = coords$Chromosome, boundpts = coords$coordinate, kb = 100000)
y <- countOverlaps(binslist100, bounds)
length(y) #27698
table(y)
prop.table(table(y))
y <- ifelse(y>=1,1,0)
prop.table(table(y))
#         0         1 
# 0.5948805 0.4051195 
mcols(binslist100)$y <- y


```

# Finding overlaps between binned genome and subcompartment intervals

```{r}

# #For Within:
# within_g <- findOverlaps(binslist, subcompartgr, type="within")
# #there are 2612294 intervals from the binned genome that overlab strictly within subcompartment intervals
# subcompoverlapwithin_g <- binslist[queryHits(within_g)]
# #attaching the subcompartment info to the within overlaps
# mcols(subcompoverlapwithin_g)$compartment <- mcols(subcompartgr)$compartment[subjectHits(within_g)]
# #adding percent overlap for each A abd B compartments
# mcols(subcompoverlapwithin_g)$percentA <- ifelse(mcols(subcompoverlapwithin_g)$compartment=="A",1,0)
# mcols(subcompoverlapwithin_g)$percentB <- ifelse(mcols(subcompoverlapwithin_g)$compartment=="B",1,0)
# 
# #For none:
# subcompoverlapnone_g <- binslist[which(binslist %outside% subcompartgr)]
# #there are 150986 intervals that have no overlaps
# #attaching the subcompartment info in the form of "N" for no overlap
# mcols(subcompoverlapnone_g)$compartment <- "N"
# #adding percent overlap
# mcols(subcompoverlapnone_g)$percentA <- 0
# mcols(subcompoverlapnone_g)$percentB <- 0
# 
# #For partial: the difference among the sets of any type of overlap and within overlaps
# subcompoverlapany_g <- findOverlaps(binslist, subcompartgr, type="any")
# partial_g <- setdiff(subcompoverlapany_g,within_g)
# #there are 5727 tad bounds that partially overlap with subcompartments
# length(unique(queryHits(partial_g))) #3034
# binslist[986] #ovelaps the following two subcompartment intervals 
# subcompartgr[c(1,136)]
# 
# 
# # Determing the percentage of partial overlaps
# 
# #determining the frequency of partial overlaps
# table(queryHits(partial_g))
# table(table(queryHits(partial_g)))
# 
# #First consider where the overlap is with only 1 subcompartment interval
# oneinterval_g <- binslist[as.numeric(rownames(as.matrix(which(table(queryHits(partial_g))==1))))]
# poneinterval_g <- findOverlapPairs(oneinterval_g, subcompartgr)
# poneinterval_g <- pintersect(poneinterval_g)
# 
# overlaps1_g <- findOverlaps(oneinterval_g, subcompartgr)
# mcols(oneinterval_g)$compartment <- mcols(subcompartgr)$compartment[subjectHits(overlaps1_g)]
# #attching compartment percent
# mcols(oneinterval_g)$percentA <- ifelse(mcols(oneinterval_g)$compartment=="A",
#                                       round(width(poneinterval_g)/width(oneinterval_g),2),
#                                       0)
# mcols(oneinterval_g)$percentB <- ifelse(mcols(oneinterval_g)$compartment=="B",
#                                       round(width(poneinterval_g)/width(oneinterval_g),2),
#                                       0)
# 
# 
# #Now we consider the overlaps between two subcompartment intervals
# twointerval_g <- binslist[as.numeric(rownames(as.matrix(which(table(queryHits(partial_g))==2))))]
# ptwointerval_g <- findOverlapPairs(twointerval_g, subcompartgr)
# ptwointerval_g <- pintersect(ptwointerval_g)
# 
# overlaps2_g <- findOverlaps(twointerval_g, subcompartgr)
# mcols(twointerval_g)$compartment <- "C"
# mcols(twointerval_g)$percentA <- round(width(ptwointerval_g[!duplicated(queryHits(overlaps2_g))])/width(twointerval_g),2)
# mcols(twointerval_g)$percentB <- round(width(ptwointerval_g[!duplicated(queryHits(overlaps2_g))])/width(twointerval_g),2)
# 
# # Concatinating all types of overlaps into 1 granges object
# #within + partial(one) + partial(two) + none = 2612294 + 341 + 2693 + 150986 = 2766314
# tad_subcomp_full <- GRangesList(subcompoverlapwithin_g, subcompoverlapnone_g, oneinterval_g, twointerval_g)
# tad_subcomp_full <- unlist(tad_subcomp_full)
# tad_subcomp_full <- tad_subcomp_full[order(substr(seqnames(tad_subcomp_full),4,5), start(tad_subcomp_full))]
# 
# table(mcols(tad_subcomp_full)$y)
# #      0       1 
# #2749544   16770
# table(mcols(tad_subcomp_full)$percentA)
# #      0     0.5       1 
# #1787171    2788  976355 
# table(mcols(tad_subcomp_full)$percentB)
# #      0     0.5       1 
# #1127436    2939 1635939
# table(mcols(tad_subcomp_full)$compartment)
# #      A       B       C       N 
# # 976450 1636185    2693  150986 
# 
# setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData")
# 
# #saving the types of overlaps
# saveRDS(tad_subcomp_full, "tad_subcomp_full.rds")
# saveRDS(subcompoverlapwithin_g, "within.rds")
# saveRDS(subcompoverlapnone_g, "none.rds")
# saveRDS(oneinterval_g, "partial1.rds")
# saveRDS(twointerval_g, "partial2.rds")

```

# Creating the dataset more model development

```{r}
#gm12878 <- data.frame(y = mcols(tad_subcomp_full)$y,
#                        A = mcols(tad_subcomp_full)$percentA,
#                        B = mcols(tad_subcomp_full)$percentB)

#omit the A and B variables created manually and instead use the subcompartment data from merlot folder

setwd("C:/Users/Spiro Stilianoudakis/Documents/TAD_data/RData/GM12878")

gm12878 <- data.frame(y = mcols(binslist1)$y)
gm12878_5kb <- data.frame(y = mcols(binslist5)$y)
gm12878_10kb <- data.frame(y = mcols(binslist10)$y)
gm12878_50kb <- data.frame(y = mcols(binslist50)$y)
gm12878_100kb <- data.frame(y = mcols(binslist100)$y)
         
# saving data frame with response vector Y for each type of genomic bin             
saveRDS(gm12878, "gm12878.rds")
saveRDS(gm12878_5kb, "gm12878_5kb.rds")
saveRDS(gm12878_10kb, "gm12878_10kb.rds")
saveRDS(gm12878_50kb, "gm12878_50kb.rds")
saveRDS(gm12878_100kb, "gm12878_100kb.rds")

# saving genomic bins
saveRDS(binslist1, "binslist1.rds")
saveRDS(binslist5, "binslist5.rds")
saveRDS(binslist10, "binslist10.rds")
saveRDS(binslist50, "binslist50.rds")
saveRDS(binslist100, "binslist100.rds")
```

